Multi-target detection with application to cryo-electron microscopy

### 2019-03-03 inverse_problems/inverse_problems.tex
plugging Will's notes (no editing yet)

\begin{proposition}
	Poisson result 
\end{proposition}
\begin{proof}
It is now easy to show how to solve for both $\gamma$ and $\sigma$ from these expressions, assuming $x$ is generic. Indeed, we observe the product
%
\begin{align}
(\gamma a_x^1) (\gamma a_x^2[1])
&=  \frac{\gamma^2}{L^2} \left( \sum_{i=0}^{L-1} x[i] \right)
\left(\sum_{j=0}^{L-2} x[j]x[j+1] \right)
\nonumber \\
&= \frac{\gamma^2}{L^2} \sum_{j=0}^{L-2} \sum_{i=0}^{L-1} x[i]x[j]x[j+1]
\nonumber \\
&= \frac{\gamma^2}{L} \left\{ \sum_{\ell=0}^{L-1} a_x^3[1,\ell]
+ \sum_{\ell=1}^{L-2} a_x^3[\ell,\ell+1]  \right\}.
\end{align}
%
Since we also observe $\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] + \sum_{\ell=1}^{L-2} \gamma a_x^3[\ell,\ell+1]$, \TODO{Is there a bias?} we can form the ratio and solve for $\gamma$:

\begin{align}
\gamma &= L \frac{(\gamma a_x^1) (\gamma a_x^2[1])}
{\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] 
	+ \sum_{\ell=1}^{L-2} \gamma a_x^3[\ell,\ell+1]}.
\end{align}
%With $\gamma$ in hand, we can estimate $\sigma^2$:
%
\begin{align}
\sigma^2 &= a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\frac{L (a^1_y)^2}{\gamma}
- (2 L - 1) (a_y^1)^2.
\end{align}

\end{proof}

@@

\TODO{Needs a complete rewrite to match the paper, in particular the modified notation and the fact we only work with autocorrelations now (not models).}

\subsubsection{First moment}

To compute the first moment of $y$, we will first condition on $M = (M_1,\dots,M_{N-L+1})$, and then average over $M$. We have:
%
\begin{align}
%
\E[y[i] | M] = \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} \E X_k^{i-j}[j]
= \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} m_y^1[j]
= \sum_{j=0}^{L-1} M_{i-j} m_y^1[j].
%
\end{align}
%
Now taking expectations over $M$ we see:
%
\begin{align}
%
\E Y[i] = \gamma \sum_{j=0}^{L-1}  m_y^1[j] = \gamma La_x^1.
%
\end{align}

@@

\subsubsection{Second moment}

Again, we will condition on $M$ first, and then take the expectation over $M$. Fix $i_1 \ne i_2$, and let $\Delta = i_2 - i_1$. Then:
%
\begin{align}
%
Y_{i_1} Y_{i_2} 
&= \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} 
\sum_{k_1=1}^{M_{i_1-j_1}}\sum_{k_2=1}^{M_{i_2-j_2}}
X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2].
%
\end{align}

We break up the double sum over $j_1$ and $j_2$ into two terms: one where $j_2 \ne j_1 + \Delta$, and one where $j_2 = j_1 + \Delta$ or equivalently $i_1-j_1 = i_2-j_2$. In the first case, all the terms are independent, and so the expectation factors. In the second case, when $k_1 \ne k_2$ we have independence, but otherwise not. This gives (all expectations are conditional on $M$):
%
\begin{align} \label{moment2-condm}
%
\E Y_{i_1} Y_{i_2}
=& \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} 
\sum_{k_1=1}^{M_{i_1-j_1}}\sum_{k_2=1}^{M_{i_2-j_2}}
\E X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2]
\nonumber \\
=& \sum_{j_1 - j_2 \ne \Delta} \sum_{k_1} 
\sum_{k_2} \E X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} \sum_{k_1 \ne k_2} 
\E X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_1 - j_1}[j_1+\Delta]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} \sum_{k_1=1}^{M_{i_1-j_1}} 
\E X_{k_1}^{i_1-j_1}[j_1] X_{k_1}^{i_1-j_1}[j_1 + \Delta] 
\nonumber \\
=& \sum_{j_1 - j_2 \ne \Delta} M_{i_1-j_1} M_{i_2 - j_2} \M_1[j_1] \M_1[j_2]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} M_{i_1-j_1}(M_{i_1-j_1} - 1) \M_1[j_1] \M_1[j_1 + \Delta]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} M_{i_1-j_1} \M_2[j_1,j_1 + \Delta] .
%
\end{align}
%
Now take expectations over the Poisson random variables, using this fact:
%
\begin{lem} \label{lem-choose}
	If $M \sim \Poisson(\gamma)$, then 
	\begin{align}
	%
	\E {M\choose k} = \frac{\gamma^k}{k!}.
	%
	\end{align}
\end{lem}

We get (now the expectation is over $M$ and $X$):
%
\begin{align}
%
\E Y_{i_1} Y_{i_2} 
=& \sum_{j_1 - j_2 \ne \Delta} \E M_{i_1-j_1} M_{i_2 - j_2} \M_1[j_1] \M_1[j_2]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} \E M_{i_1-j_1}(M_{i_1-j_1} - 1) \M_1[j_1] \M_1[j_1 + \Delta]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} \E M_{i_1-j_1} \M_2[j_1,j_1 + \Delta]
\nonumber \\
=& \sum_{j_1 - j_2 \ne \Delta} \gamma^2 \M_1[j_1] \M_1[j_2]
+ \sum_{j_1 = 0}^{L-1} \gamma^2 \M_1[j_1] \M_1[j_1 + \Delta]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} \gamma \M_2[j_1,j_1 + \Delta]
\nonumber \\
=&  \bigg(\gamma \sum_{j = 0}^{L-1} \M_1[j] \bigg)^2
+ \gamma \sum_{j = 0}^{L-1} \M_2[j,j + \Delta]
\nonumber \\
=&  (\gamma \L_1)^2 + \gamma \L_2(\Delta).
%
\end{align}

But the first term in the sum is just the square of the first moment of $Y$; so from the first two moments we can recover $\gamma \L_2(\Delta)$, which is just the expected power spectrum of the random vector $X$, i.e.\ the usual second moment we have been working with.

@@

\subsubsection{Third moment}

For three distinct $i_1$, $i_2$ and $i_3$, we let $\Delta_1 = i_2 - i_1$ and $\Delta_2 = i_3 - i_1$. We have:
%
\begin{align}
%
&Y_{i_1} Y_{i_2} Y_{i_3}
\nonumber \\
&= \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} \sum_{j_3=0}^{L-1} 
\sum_{k_1=1}^{M_{i_1-j_1}}\sum_{k_2=1}^{M_{i_2-j_2}} \sum_{k_3=1}^{M_{i_3-j_3}}
X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2] X_{k_3}^{i_3 - j_3}[j_3].
\end{align}
%
We will break up the outer three sums into disjoint sums with the following ranges of indices:
%
\begin{enumerate}

@@

For Case \ref{case1}, we have $\ell \equiv i_1 - j_1 = i_2 - j_2 = i_3 - j_3$. We further break up the sum:
%
\begin{align}
%
&\sum_{j=0}^{L-1} \sum_{k_1=1}^{M_\ell} \sum_{k_2=1}^{M_\ell} \sum_{k_3=1}^{M_\ell} 
X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
\nonumber \\
=& \underbrace{ \sum_{j=0}^{L-1} \sum_{k_i \text{distinct}} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(a)}}
\nonumber \\
&+\underbrace{ \sum_{j=0}^{L-1} \sum_{k_1=k_2\ne k_3} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(b)}}
\nonumber \\
&+\underbrace{ \sum_{j=0}^{L-1} \sum_{k_1=k_3\ne k_2} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(c)}}
\nonumber \\
&+\underbrace{ \sum_{j=0}^{L-1} \sum_{k_2=k_3\ne k_1} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(d)}}
\nonumber \\
&+\underbrace{ \sum_{j=0}^{L-1} \sum_{k_1=k_2=k_3} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(e)}}.
%
\end{align}

For term (a), the expectation conditional on $M$ is:
%
\begin{align}
%
\sum_{j=0}^{L-1} M_\ell(M_\ell-1)(M_\ell-2)\M[j] \M[j+\Delta_1] \M[j+\Delta_2].
%
\end{align}
%
Using Lemma \ref{lem-choose}, the unconditional expectation of (a) is then:
%
\begin{align} \label{aaaa}
%
\gamma^3 \sum_{j=0}^{L-1} \M_1[j] \M_1[j+\Delta_1] \M_1[j+\Delta_2].
%
\end{align}

For term (b), the expectation conditional on $M$ is:
%
\begin{align}
%
\sum_{j=0}^{L-1} M_\ell (M_\ell - 1) \M_2[j,j+\Delta_1] \M_1[j + \Delta_2]
%
\end{align}
%
and then again using Lemma \ref{lem-choose} we get the expected value:
%
\begin{align} \label{bbbb}
%
\gamma^2 \sum_{j=0}^{L-1} \M_2[j,j+\Delta_1] \M_1[j+\Delta_2].
%
\end{align}

Similarly, the expected values of terms (c) and (d) are:
%
\begin{align} \label{cccc}
%
\gamma^2 \sum_{j=0}^{L-1} \M_2[j,j+\Delta_2] \M_1[j+\Delta_1].
%
\end{align}
%
and
%
\begin{align} \label{dddd}
%
\gamma^2 \sum_{j=0}^{L-1} \M_2[j+\Delta_1,j+\Delta_2] \M_1[j].
%
\end{align}

Finally, the expected value of term (e) is easily shown to be:
%
\begin{align} \label{eeee}
%
\gamma \sum_{j=0}^{L-1} \M_3[j,j+\Delta_1,j+\Delta_2].
%
\end{align}
%
This concludes the computation for Case \ref{case1}.

Moving onto Case \ref{case2}, we have $\ell_1 \equiv i_1 - j_1 = i_2 - j_2$, and also define $\ell_2 \equiv i_3 - j_3$. By definition, $\ell_1 \ne \ell_2$. The sum is:
%
\begin{align}
%
& \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2}
\sum_{1 \le k_1,k_2 \le M_{\ell_1}} \sum_{k_3=1}^{M_{\ell_2}}
X_{k_1}^{\ell_1}[j_1] X_{k_2}^{\ell_1}[j_1 + \Delta_1] X_{k_3}^{\ell_2}[j_3]   
\nonumber \\
=& \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2} \sum_{k_3=1}^{M_{\ell_2}}
\Bigg\{
\sum_{1 \le k_1 \ne k_2 \le M_{\ell_1}} 
X_{k_1}^{\ell_1}[j_1] X_{k_2}^{\ell_1}[j_1 + \Delta_1] X_{k_3}^{\ell_2}[j_3]
\nonumber \\
&       + \sum_{k_1=1}^{M_{\ell_1}} X_{k_1}^{\ell_1}[j_1] X_{k_1}^{\ell_1}[j_1 + \Delta_1] 
X_{k_3}^{\ell_2}[j_3]  \Bigg\}.
%
\end{align}
%
Taking expectations conditional on $M$, we then get:
%
\begin{align}
%
& \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2} 
\Bigg(M_{\ell_1} (M_{\ell_1}-1) M_{\ell_2} \M_1[j_1] \M_1[j_1 + \Delta_1] \M_1[j_3]
\nonumber \\
&       + M_{\ell_1} M_{\ell_2} \M_2[j_1,j_1+\Delta_1] \M_1[j_3] \Bigg).
%
\end{align}

Taking expectations over $M$ and using Lemma \ref{lem-choose} then gives:
%
\begin{align}
%
& \gamma^3 \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2} 
\M_1[j_1] \M_1[j_1 + \Delta_1] \M_1[j_3]
\label{ffff} \\
& + \gamma^2 \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2}  
\M_2[j_1,j_1+\Delta_1] \M_1[j_3].
\label{gggg}
%
\end{align}

Similarly, Cases \ref{case3} and \ref{case4} give the expressions:
%
\begin{align}
%
& \gamma^3 \sum_{j_1=0}^{L-1} \sum_{j_2 \ne j_1 + \Delta_1} 
\M_1[j_1] \M_1[j_1 + \Delta_2] \M_1[j_2]
\label{hhhh} \\
& + \gamma^2 \sum_{j_1=0}^{L-1} \sum_{j_2 \ne j_1 + \Delta_1}  
\M_2[j_1,j_1+\Delta_2] \M_1[j_2]
\label{iiii}
%
\end{align}
%
and
%
\begin{align}
%
& \gamma^3 \sum_{j_2=0}^{L-1} \sum_{j_1 \ne j_2} 
\M_1[j_1] \M_1[j_2 + \Delta_1] \M_1[j_2 + \Delta_2]
\label{jjjj}\\
& + \gamma^2 \sum_{j_2=0}^{L-1} \sum_{j_1 \ne j_2}  
\M_2[j_2+\Delta_1,j_2+\Delta_2] \M_1[j_1].
\label{kkkk}
%
\end{align}

Finally, in Case \ref{case5} we have $i_1 - j_1$, $i_2 - j_2$, and $i_3 - j_3$ are all pairwise distinct. Consequently, the $X$ variables are always independent, and the expectation conditional on $M$ (letting $\ell_q = i_q - j_q$, $q=1,2,3$),
%
\begin{align}
%
\sum_{j_1,j_2,j_3} M_{\ell_1} M_{\ell_2} M_{\ell_3} \M_1[j_1] \M_1[j_2] \M_1[j_3];
%
\end{align}
%
since the $M_{\ell_q}$'s are pairwise independent, $q=1,2,3$, the expectation over $M$ then yields:
%
\begin{align} \label{llll}
%
\gamma^3 \sum_{j_1,j_2,j_3} \M_1[j_1] \M_1[j_2] \M_1[j_3].
%
\end{align}

Now we add all the terms from Cases \ref{case1} to \ref{case5}. Expressions \eqref{aaaa}, \eqref{ffff}, \eqref{hhhh}, \eqref{jjjj}, and \eqref{llll} sum to the expression:
%
\begin{align}
%
(\gamma \L_1)^3.
%
\end{align}

Note that this is obtained directly from the first moment. Expressions \eqref{bbbb}, \eqref{cccc}, \eqref{dddd}, \eqref{gggg},\eqref{iiii}, and \eqref{kkkk} sum to the expression:
%
\begin{align}
%
\gamma \L_1  \cdot 
( \gamma\L_2(\Delta_1) + \gamma\L_2(\Delta_2) + \gamma\L_2(\Delta_2-\Delta_1)).
%
\end{align}
%
Again, note that this is obtained directly from the first two moments. Finally, expression \eqref{eeee} is simply:
%
\begin{align}
%
\gamma \L_3(\Delta_1,\Delta_2)
%
\end{align}
%
which is the usual third-order auto-correlation.

>>>>

The proof is provided in Section~\ref{sec:proof_uniqueness_poisson}.
\begin{proposition} \label{prop:uniqueness_poisson}
	Poisson result 
\end{proposition}

@@

Throughout this section, we will use the capital letter $X$ to represent the random unobserved vector of length $L$, the capital letter $Y$ to represent the random observed vector of length $N$, and the capital letter $S$ to represent the unobserved vector of occurrences.

We will denote by $\M_l$ the moment tensors of $X$:
%
\begin{align}
%
\M_1[i] = \E X[i], \quad 0 \le i \le L-1,
%
\end{align}
%
\begin{align}
%
\M_2[i,j] = \E X[i] X[j], \quad 0 \le i,j \le L-1,
%
\end{align}
%
\begin{align}
%
\M_3[i,j,k] = \E X[i] X[j] X[k], \quad 0 \le i,j,k \le L-1.
%
\end{align}

We obtain the autocorrelations $a_x^l$ of $x$ by averaging over a slice of the moment tensors:
%
\begin{align}
%
a_x^1 = \frac{1}{L}\sum_{i=0}^{L-1} \M_1[i],
%
\end{align}
%
\begin{align}
%
a_x^2[\Delta] = \frac{1}{L} \sum_{i=0}^{L-1} \M_2[i,i+\Delta],
%
\end{align}
%
and
%
\begin{align}
%
a_x^3[\Delta_1,\Delta_2] = \frac{1}{L}\sum_{i=0}^{L-1} \M_3[i,i+\Delta_1,i+\Delta_2].
%
\end{align}

We will make repeated use of the following elementary lemma:
%
\begin{lem} \label{lem-choose}
	%
	If $M \sim \Poisson(\lambda)$, then 
	\begin{align}
	%
	\E {M\choose k} = \frac{\lambda^k}{k!}.
	%
	\end{align}
\end{lem}

@@

\subsubsection{Computing $a_y^1$}

We will first condition on the vector $s$ of locations of the subsignals in $y$, and then average over $s$. We will denote by $X_1^{i},\dots,X_{S[i]}^i$ the random vectors with position 0 located at $i$. We have:
%
\begin{align}
%
\E[ Y[i] | S ] = \sum_{j=0}^{L-1} \sum_{k=1}^{S[i-j]} \E X_k^{i-j}[j]
= \sum_{k=1}^{S[i-j]} L a_x^1
= S[i-j] L a_x^1.
%
\end{align}
%
Now taking expectations over $s$ and using $\E S[i-j] = \gamma/L$ we get:
%
\begin{align}
%
\E Y[i] = \gamma a_x^1.
%
\end{align}
%
Consequently,
%
\begin{align}
a_y^1 = \frac{1}{N} \sum_{i=1}^n \E Y[i] = \gamma a_x^1.
\end{align}

@@

\subsubsection{Computing $a_y^2$}

First we will consider the noise-free case, where $\sigma = 0$. We will condition on $s$ first, and then take the expectation over $s$. Fix $i_1 \ne i_2$, and let $\Delta = i_2 - i_1$. Then:
%
\begin{align}
%
Y[i_1] Y[i_2]
&= \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} 
\sum_{k_1=1}^{S[i_1-j_1]}\sum_{k_2=1}^{S[i_2-j_2]}
X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2].
%
\end{align}

We break up the double sum over $j_1$ and $j_2$ into two terms: one where $j_2 \ne j_1 + \Delta$, and one where $j_2 = j_1 + \Delta$ or equivalently $i_1-j_1 = i_2-j_2$. In the first case, all the terms are independent, and so the expectation factors. In the second case, when $k_1 \ne k_2$ we have independence, but otherwise not. This gives (all expectations are conditional on $s$):
%
\begin{align} \label{moment2-condm}
%
\E Y[i_1] Y[i_2]
=& \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} 
\sum_{k_1=1}^{S[i_1-j_1]}\sum_{k_2=1}^{S[i_2-j_2]}
\E X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2]
\nonumber \\
=& \sum_{j_1 - j_2 \ne \Delta} \sum_{k_1} 
\sum_{k_2} \E X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} \sum_{k_1 \ne k_2} 
\E X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_1 - j_1}[j_1+\Delta]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} \sum_{k_1=1}^{S[i_1-j_1]} 
\E X_{k_1}^{i_1-j_1}[j_1] X_{k_1}^{i_1-j_1}[j_1 + \Delta] 
\nonumber \\
=& \sum_{j_1 - j_2 \ne \Delta} S[i_1-j_1] S[i_2 - j_2] \M_1[j_1] \M_1[j_2]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} S[i_1-j_1](S[i_1-j_1] - 1) \M_1[j_1] \M_1[j_1 + \Delta]
\nonumber \\
& + \sum_{j_1 = 0}^{L-1} S[i_1-j_1] \M_2[j_1,j_1 + \Delta] .
%
\end{align}
%
Now take expectations over the Poisson random variables, using Lemma \ref{lem-choose}:
%
\begin{align}
%
\E Y[i_1] Y[i_2]
=& \sum_{j_1 - j_2 \ne \Delta} \E S[i_1-j_1] S[i_2 - j_2] \M_1[j_1] \M_1[j_2]
\nonumber \\
&       + \sum_{j_1 = 0}^{L-1} \E S[i_1-j_1](S[i_1-j_1] - 1) \M_1[j_1] \M_1[j_1 + \Delta]
\nonumber \\
&       + \sum_{j_1 = 0}^{L-1} \E S[i_1-j_1] \M_2[j_1,j_1 + \Delta]
\nonumber \\
=& \frac{1}{L^2}\sum_{j_1 - j_2 \ne \Delta} \gamma^2 \M_1[j_1] \M_1[j_2]
+ \frac{1}{L^2} \sum_{j_1 = 0}^{L-1} \gamma^2 \M_1[j_1] \M_1[j_1 + \Delta]
\nonumber \\
&       + \frac{1}{L}\sum_{j_1 = 0}^{L-1} \gamma \M_2[j_1,j_1 + \Delta]
\nonumber \\
=&  \bigg(\frac{\gamma}{L}  \sum_{j = 0}^{L-1} \M_1[j] \bigg)^2
+ \frac{\gamma}{L} \sum_{j = 0}^{L-1} \M_2[j,j + \Delta]
\nonumber \\
=&  (\gamma a_x^1)^2 + \gamma a_x^2[\Delta].
%
\end{align}
%
For positive $\sigma$, we observe that any terms linear in the noise vanish in expectation. Denoting by $X^*$ the clean signal component of length $N$, so $Y = X^* + \ep$, we have:
%
%
\begin{align}
\E Y[i_1] Y[i_2] = \E X^*[i_1] X^*[i_2] + \E \ep[i_1] \ep[i_2]
= (\gamma a_x^1)^2 + \gamma a_x^2[\Delta] + \sigma^2 \delta[i_1 - i_2].
\end{align}
%
We conclude by averaging over $i_1$ and $i_2$, leaving the gap $\Delta = i_2 - i_1$ fixed.

\subsubsection{Computing $a_x^3$}

We will first assume $\sigma = 0$. For three distinct $i_1$, $i_2$ and $i_3$, we let $\Delta_1 = i_2 - i_1$ and $\Delta_2 = i_3 - i_1$. We have:
%
\begin{align}
%
& Y[i_1] Y[i_2] Y[i_3]
= \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} \sum_{j_3=0}^{L-1} 
\sum_{k_1=1}^{S[i_1-j_1]}\sum_{k_2=1}^{S[i_2-j_2]} \sum_{k_3=1}^{S[i_3-j_3]}
X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2] X_{k_3}^{i_3 - j_3}[j_3].
\end{align}
%
We will break up the outer three sums into disjoint sums with the following ranges of indices:
%
\begin{enumerate}

@@

For Case \ref{case1}, we have $\ell \equiv i_1 - j_1 = i_2 - j_2 = i_3 - j_3$. We further break up the sum:
%
\begin{align}
%
&\sum_{j=0}^{L-1} \sum_{k_1=1}^{S[\ell]} \sum_{k_2=1}^{S[\ell]} \sum_{k_3=1}^{S[\ell]} 
X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
\nonumber \\
=& \underbrace{ \sum_{j=0}^{L-1} \sum_{k_i \text{distinct}} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(a)}}
\nonumber \\
&+\underbrace{ \sum_{j=0}^{L-1} \sum_{k_1=k_2\ne k_3} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(b)}}
\nonumber \\
&+\underbrace{ \sum_{j=0}^{L-1} \sum_{k_1=k_3\ne k_2} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(c)}}
\nonumber \\
&+\underbrace{ \sum_{j=0}^{L-1} \sum_{k_2=k_3\ne k_1} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(d)}}
\nonumber \\
&+\underbrace{ \sum_{j=0}^{L-1} \sum_{k_1=k_2=k_3} 
	X_{k_1}^{\ell}[j] X_{k_2}^{\ell}[j + \Delta_1] X_{k_3}^{\ell}[j + \Delta_2]
}_{\text{(e)}}.
%
\end{align}

For term (a), the expectation conditional on $S$ is:
%
\begin{align}
%
\sum_{j=0}^{L-1} S[\ell](S[\ell]-1)(S[\ell]-2)\M[j] \M[j+\Delta_1] \M[j+\Delta_2].
%
\end{align}
%
Using Lemma \ref{lem-choose}, the unconditional expectation of (a) is then:
%
\begin{align} \label{aaaa}
%
\frac{\gamma^3}{L^3} \sum_{j=0}^{L-1} \M_1[j] \M_1[j+\Delta_1] \M_1[j+\Delta_2].
%
\end{align}

For term (b), the expectation conditional on $S$ is:
%
\begin{align}
%
\sum_{j=0}^{L-1} S[\ell] (S[\ell] - 1) \M_2[j,j+\Delta_1] \M_1[j + \Delta_2]
%
\end{align}
%
and then again using Lemma \ref{lem-choose} we get the expected value:
%
\begin{align} \label{bbbb}
%
\frac{\gamma^2}{L^2} \sum_{j=0}^{L-1} \M_2[j,j+\Delta_1] \M_1[j+\Delta_2].
%
\end{align}

Similarly, the expected values of terms (c) and (d) are:
%
\begin{align} \label{cccc}
%
\frac{\gamma^2}{L^2} \sum_{j=0}^{L-1} \M_2[j,j+\Delta_2] \M_1[j+\Delta_1].
%
\end{align}
%
and
%
\begin{align} \label{dddd}
%
\frac{\gamma^2}{L^2} \sum_{j=0}^{L-1} \M_2[j+\Delta_1,j+\Delta_2] \M_1[j].
%
\end{align}

Finally, the expected value of term (e) is easily shown to be:
%
\begin{align} \label{eeee}
%
\frac{\gamma}{L} \sum_{j=0}^{L-1} \M_3[j,j+\Delta_1,j+\Delta_2].
%
\end{align}
%
This concludes the computation for Case \ref{case1}.

Moving onto Case \ref{case2}, we have $\ell_1 \equiv i_1 - j_1 = i_2 - j_2$, and also define $\ell_2 \equiv i_3 - j_3$. By definition, $\ell_1 \ne \ell_2$. The sum is:
%
\begin{align}
%
& \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2}
\sum_{1 \le k_1,k_2 \le S[\ell_1]} \sum_{k_3=1}^{S[\ell_2]}
X_{k_1}^{\ell_1}[j_1] X_{k_2}^{\ell_1}[j_1 + \Delta_1] X_{k_3}^{\ell_2}[j_3]   
\nonumber \\
=& \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2} \sum_{k_3=1}^{S[\ell_2]}
\Bigg\{  \sum_{1 \le k_1 \ne k_2 \le S[\ell_1]} 
X_{k_1}^{\ell_1}[j_1] X_{k_2}^{\ell_1}[j_1 + \Delta_1] X_{k_3}^{\ell_2}[j_3]
\nonumber \\
& + \sum_{k_1=1}^{S[\ell_1]} X_{k_1}^{\ell_1}[j_1] X_{k_1}^{\ell_1}[j_1 + \Delta_1] 
X_{k_3}^{\ell_2}[j_3]  \Bigg\}.
%
\end{align}
%
Taking expectations conditional on $S$, we then get:
%
\begin{align}
%
\sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2} 
\Bigg( & S[\ell_1] (S[\ell_1]-1) S[\ell_2] \M_1[j_1] \M_1[j_1 + \Delta_1] \M_1[j_3]
\nonumber \\
&  + S[\ell_1] S[\ell_2] \M_2[j_1,j_1+\Delta_1] \M_1[j_3] \Bigg).
%
\end{align}

Taking expectations over $S$ and using Lemma \ref{lem-choose} then gives:
%
\begin{align}
%
& \frac{\gamma^3}{L^3} \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2} 
\M_1[j_1] \M_1[j_1 + \Delta_1] \M_1[j_3]
\label{ffff} \\
& + \frac{\gamma^2}{L^2} \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \Delta_2}  
\M_2[j_1,j_1+\Delta_1] \M_1[j_3].
\label{gggg}
%
\end{align}

Similarly, Cases \ref{case3} and \ref{case4} give the expressions:
%
\begin{align}
%
& \frac{\gamma^3}{L^3} \sum_{j_1=0}^{L-1} \sum_{j_2 \ne j_1 + \Delta_1} 
\M_1[j_1] \M_1[j_1 + \Delta_2] \M_1[j_2]
\label{hhhh} \\
& + \frac{\gamma^2}{L^2} \sum_{j_1=0}^{L-1} \sum_{j_2 \ne j_1 + \Delta_1}  
\M_2[j_1,j_1+\Delta_2] \M_1[j_2]
\label{iiii}
%
\end{align}
%
and
%
\begin{align}
%
& \frac{\gamma^3}{L^3} \sum_{j_2=0}^{L-1} \sum_{j_1 \ne j_2} 
\M_1[j_1] \M_1[j_2 + \Delta_1] \M_1[j_2 + \Delta_2]
\label{jjjj}\\
& + \frac{\gamma^2}{L^2} \sum_{j_2=0}^{L-1} \sum_{j_1 \ne j_2}  
\M_2[j_2+\Delta_1,j_2+\Delta_2] \M_1[j_1].
\label{kkkk}
%
\end{align}

Finally, in Case \ref{case5} we have $i_1 - j_1$, $i_2 - j_2$, and $i_3 - j_3$ are all pairwise distinct. Consequently, the $X$ variables are always independent, and the expectation conditional on $S$ (letting $\ell_q = i_q - j_q$, $q=1,2,3$),
%
\begin{align}
%
\sum_{j_1,j_2,j_3} S[\ell_1] S[\ell_2] S[\ell_3] \M_1[j_1] \M_1[j_2] \M_1[j_3];
%
\end{align}
%
since the $S[\ell_q]$'s are pairwise independent, $q=1,2,3$, the expectation over $S$ then yields:
%
\begin{align} \label{llll}
%
\frac{\gamma^3}{L^3} \sum_{j_1,j_2,j_3} \M_1[j_1] \M_1[j_2] \M_1[j_3].
%
\end{align}

Now we add all the terms from Cases \ref{case1} to \ref{case5}. Expressions \eqref{aaaa}, \eqref{ffff}, \eqref{hhhh}, \eqref{jjjj}, and \eqref{llll} sum to the expression:
%
\begin{align}
%
(\gamma a_x^1)^3.
%
\end{align}

Expressions \eqref{bbbb}, \eqref{cccc}, \eqref{dddd}, \eqref{gggg},\eqref{iiii}, and \eqref{kkkk} sum to the expression:
%
\begin{align}
%
\gamma a_x^1  \cdot 
( \gamma a_x^2[\Delta_1] + \gamma a_x^2[\Delta_2] + \gamma a_x^2[\Delta_2-\Delta_1]).
%
\end{align}
%
Finally, expression \eqref{eeee} is simply:
%
\begin{align}
%
\gamma a_x^3[\Delta_1,\Delta_2].
%
\end{align}
%

Now when $\sigma > 0$, we write $Y = X^* + \ep$, and
%
\begin{align}
\E Y[i_1] Y[i_2] Y[i_3] =& \E X^*[i_1] X^*[i_2] X^*[i_3] + \E X^*[i_1] \ep[i_2] \ep[i_3]
\nonumber \\
& + \E\ep[i_1] X^*[i_2] \ep[i_3] + \E \ep[i_1] \ep[i_2] X^*[i_3] 
+ \E \ep[i_1] \ep[i_2]\ep[i_3]
\nonumber \\
=& \E X^*[i_1] X^*[i_2] X^*[i_3] + \gamma  a_x^1 \cdot \sigma^2 \delta[i_2-i_3]
\nonumber \\
& + \gamma a_x^1 \cdot \sigma^2 \delta[i_3-i_1] 
+ \gamma a_x^1 \cdot \sigma^2 \delta[i_2-i_1] .
\end{align}
